---
title: "us-renewables"
author: "Tyler Cobian"
date: "2/6/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewables)
```{r}
us_renew<- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

- Make the descriptions all lowercase
- Only keep observations where "desription" variable contains "consumption"
- Remove any observation where "description" variable contains "total"
```{r}
us_renew_clean<- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))
```

### Convert `yyyymm` column as a date
```{r}
renew_date<- us_renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)
# make a version where the month and year are in seperate columns
renew_pased<- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at the new data:
```{r}
renew_gg<- ggplot(renew_date, aes(x = month_sep, 
                                  y = value, 
                                  group = description,  
                                  color = description))+
  geom_line()
renew_gg
```

Updating colors with paletter palettes:
```{r}
# Look here for adding new color palettes to graphs
# in scale_color_paletter_d the "d" stands for discrete you can put "c" and get continuous palettes
# type this in View(palettes_d_names) to view all the palettes
renew_gg+
  scale_color_paletteer_d("calecopal::bigsur")+
  theme_minimal()
```

Coaerce the renew_parse to a tsibble
```{r}
renew_ts<- as_tsibble(renew_pased, key = description, index = month_sep)
```

Look at the new data in a couple different ways
```{r}
renew_ts %>% 
  autoplot(value)
renew_ts %>% 
  gg_subseries(value)
# make a graph of season 
ggplot(renew_pased, aes(x = month,
                        y = value,
                        group = year))+
  geom_line(aes(color = year))+
  facet_wrap(~description,
             ncol = 1,
             scales = "free",
             strip.position = "right")
```

```{r}
hydro_ts<- renew_ts %>% 
  filter(description == "hydroelectric power consumption")
hydro_ts %>% 
  autoplot(value)
hydro_ts %>% 
  gg_subseries(value)

ggplot(hydro_ts, aes(x = year, y = value, group = year))+
  geom_line(aes(color = year))
```

What is I want the quarterly average consumption for hydro
```{r}
hydro_quarterly<- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarise(avg_consumption = mean(value))
head(hydro_quarterly)
```

### decompose the hydro_ts data
```{r}
dcmp<- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% 
  autoplot()
```

```{r}
hist(components(dcmp)$remainder)
```

Now lets look at the ACF:
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

#Model (if you want to do this on your own beware)
```{r}
hydro_model<- hydro_ts %>% 
  model(
    ARIMA(value), # ARIMA model
    ETS(value) # ETS model (ignores seasonality)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% 
  autoplot(filter(hydro_ts, year(month_sep)>2010))
```

### Make a world map
```{r}
world <- read_sf(dsn = here("data","TM_WORLD_BORDERS_SIMPL-0.3-1"), layer = "TM_WORLD_BORDERS_SIMPL-0.3") %>% clean_names()

# Quick & easy option to see those polygons (also for points, lines!)
mapview(world)

```













